Revisiting divisible, injective and flabby abelian groups from a constructive point of view

### 2022-09-27 slides-fischbachau2022.tex
¬¬-motivation finished

\begin{frame}{The double-negation embedding}
  \textbf{Thm.} Every infinite sequence~$\alpha : \NN \to \NN$ is \emph{good}
  in that there are numbers~$i < j$ such that~$\alpha(i) \leq \alpha(j)$.

  \textbf{Thm.} Every infinite binary sequence~$\alpha : \NN \to \mathbbm{2}$
  contains repeating entries.

  {\emph{Proof.} By the \bad{infinite box principle}, which follows
  from~\badbox{\textsc{lem}}, the sequence contains infinitely many zeros or
  infinitely many ones. In either case the claim follows.\par}

  How to extract \good{constructive proofs} from these
  \bad{classical proofs}?
  \pause

@@

  \textbf{Thm.} Let~$X$ be an inhabited set of natural numbers. Then~$X$
  contains a minimal element.

  {\emph{Proof.} By~\badbox{\textsc{lem}}, \ldots\par}
\end{frame}

@@

>>>>

\begin{frame}{The double-negation embedding}
  \vspace*{-0.8em}
  \mbox{How to extract \good{constructive proofs} from
  \bad{classical proofs} as the following?}

  \begin{enumerate}
    \item \textbf{Thm.} Every infinite sequence~$\alpha : \NN \to \NN$ is \emph{good}
    in that there are numbers~$i < j$ such that~$\alpha(i) \leq \alpha(j)$.
    \medskip

    {\emph{Proof.} By~\badbox{\textsc{lem}}, there is a minimal value~$\alpha(i)$.
    Set~$j \defeq i + 1$.\par}
    \medskip
    \pause

    \item \textbf{Thm.} Every infinite binary sequence
    contains repeated terms.
    \medskip

    {\emph{Proof.} By the \bad{infinite box principle},
    infinitely many terms are zeros or are ones. In either case the claim
    follows.\par}
    \pause
  \end{enumerate}
  \vspace*{-0.7em}

@@

  \textbf{Thm.} Let~$X$ be an inhabited set of natural numbers.
  \begin{enumerate}
    \item If~\badbox{\textsc{lem}}, then~$X$ contains a minimal element.
    \item<5-> If~$X$ is \hil{detachable}, then~$X$ contains a minimal element.
    \item<7-> It is \notnot the case that~$X$ contains a minimal element.
  \end{enumerate}

  \justifying
  \only<1-3>{\ \\\ \\\ \\\ }
  \only<4-5>{\emph{Proof of~\normalnumber{1}.} There is some~$n \in X$.
  By~\badbox{\textsc{lem}}, either~$\exists k \in X\_ k < n$ \\ or not. In the
  first case, we continue by induction. Else~$n$ is minimal.\\\ \par}
  \only<6-7>{\mbox{\emph{Proof of~\normalnumber{2}.} There is some~$n \in X$.
  By~\good{assumption}, either~$\exists k \in X\_ k < n$} or not. In the
  first case, we continue by induction. Else~$n$ is minimal.\\\ \par}
  \only<8>{\emph{Proof of~\normalnumber{3}.} There is some~$n \in X$. Assume that~$X$ does not
  contain a minimum. Then it is not the case that~$\exists k \in X\_ k <
  n$, as else~$\bot$ by induction. Hence~$n$ is minimal. This is a
  contradiction.\par}
\end{frame}

\begin{frame}
  \centering\bigskip
  \includegraphics[height=9.5em]{phantoms}
  \medskip

  \large Lecture III: \\
  \hil{Extracting constructive proofs from invalid* proofs}

  \normalsize
\end{frame}

@@

Claim: (∃n ∈ X) ⇒ X has a minimum.

Classical proof:
By LEM, ∃(k ∈ X). k < n or not.
In the first case, we conclude by induction.
In the second case, we conclude that n is minimal:
  Given k ∈ X, we have k < n or n ≤ k.
  In the first case, ⊥ and hence conclusion.
  In the second case, good.

Translated proof:
Assume that X does not have a minimum, show ⊥.
May conclude that ¬(∃(k ∈ X). k < n). For else, X inhabited by <n, so ⊥ by induction.
Thus n is a minimum:
  Given k ∈ X, we have k < n or n ≤ k.
  In the first case, ⊥ and hence conclusion.
  In the second case, good.
